Compositional Certified Resource Bounds\ifshort{\vspace{-1.15cm}}{}

### 2015-03-16 paper/poly.tex
more proofs, the notations are clumsy

Let $V$ be a set of variabls.  An index $I \in \ind(V)$ denotes a
\emph{base polynomial} $P_I : \State \to \Q$ that maps a program
state $\state$ to a rational number.  The definition of $P_I$ is
$
P_I(\state) = \prod_{(x,y)} |[\state(x),\state(y)]|^{i_{(x,y)}}.
$
Remark that following the standard convention for $\prod$, the
empty index represents the constant base polynomial 1.  We will
later use the notation $P_{(x,y)}(\state) = |[\state(x),\state(y)]|$
to define $P_I$ as $P_I = \prod_{(x,y)} P_{(x,y)}$.

@@

The substitution operation $I[\square \gets x]$ defines an index $I'$
in $\ind(V)$ with the same degree as $I$.  This index satisfies
$$
i'_{(x,z)} {=} i_{(x,z)} {+} i_{(\square,z)}, \;
i'_{(z,x)} {=} i_{(z,x)} {+} i_{(z,\square)}, \;
i'_{(u,v)} {=} i_{(u,v)} \mbox{ o.w.}
$$
Notations may seem heavy, but what they really express is that
replacing $\square$ by $x$ in $x^2\square^3z^2$ yields $x^5z^2$.

We now define an interpretation $P_.^y$ of indices containing
$\square$:
\begin{align*}
P^y_{(\square,z)}(\state) &=
  \left| [\sigma(x){+}\sigma(y), \sigma(z)] \,\triangle\,
         [\sigma(x), \sigma(z)] \right| \\
P^y_{(z,\square)}(\state) &=
  \left| [\sigma(z), \sigma(x){+}\sigma(y)] \,\triangle\,
         [\sigma(z), \sigma(x)] \right| \\
P^y_{(\square, \square)}(\state) &= P_{(x,\square)}(\state)
  = P_{(\square,x)}(\state) = 0 \\
P^y_{(u,v)}(\state) &= |[\sigma(u), \sigma(v)]|,
\end{align*}
we use $A \,\triangle\, B$ to denote the symmetric difference
of $A$ and $B$.
The quantity $P^y_{(\square,z)}(\state)$ is the size change of
the interval $[x, z]$.  If $\state$ is the state before the
assignment and $\state'$ is the state after the assignment, we
have
$P_{(x,z)}(\state) = P^y_{(x,z)}(\state') + P^y_{(\square,z)}(\state')$
and
$P_{(z,x)}(\state) = P^y_{(z,x)}(\state') - P^y_{(z,\square)}(\state')$.
With these definitions, we have the following lemma.
\begin{lemma}
If $\state \rightarrow_C \state'$ and $\state(y)>0$ then
$P_Q(\state) = P^y_{\qsh^+_x Q}(\state')$.
\end{lemma}
\begin{proof}
\small
\begin{align*}
&P^y_{\qsh_x^+ Q}(\state')
= \sum_I q'_I \prod_{(u,v)} P_{uv}^{uv} \\
&= \sum_I q_{I[\square\gets x]} P_I \prod_z
  \binom{\square z {+} xz}{\square z}P_{\square z}^{\square z} P_{xz}^{xz}
  \binom{z\! \square\! {+} zx}{z \square}(-P_{z \square})^{z \square} P_{zx}^{zx} \\
&= \sum_J q_J P_J \sum_{I[\square\gets x] = J} \prod_z
  \binom{\square z {+} xz}{\square z}P_{\square z}^{\square z} P_{xz}^{xz}
  \binom{z\! \square\! {+} zx}{z \square}(-P_{z \square})^{z \square} P_{zx}^{zx} \\
&= \sum_J q_J P_J \prod_z
  \left( \sum_{\square z + xz = j_{xz}}
    \binom{\square z {+} xz}{\square z}P_{\square z}^{\square z} P_{xz}^{xz} \right)
  \left( \sum_{z\square {+} zx = j_{zx}} \dots \right) \\
&= \sum_J q_J P_J \prod_z
  (P_{xz} + P_{\square z})^{j_{xz}}
  (P_{zx} - P_{z \square})^{j_{zx}} \\
&= \sum_J q_J P_J(\state) = P_Q(\state)
\end{align*}
\end{proof}

@@

>>>>

Let $V$ be a set of variabls.  An index $I \in \ind(V)$ denotes a
\emph{base polynomial} $P_I : \State \to \Q$ that maps a program
state $\state$ to a rational number.  The definition of $P_I$ is
$
P_I(\state) = \prod_{(x,y)} |[\state(x),\state(y)]|^{i_{(x,y)}}.
$
Remark that following the standard convention for $\prod$, the
empty index represents the constant base polynomial 1.  We will
later use the notation $P_{xy}(\state) = |[\state(x),\state(y)]|$
to define $P_I$ as $P_I = \prod_{(x,y)} P_{xy}$.

@@

The substitution operation $I[\square \gets x]$ defines an index $J$
in $\ind(V)$ with the same degree as $I$.  This index satisfies
$$
j_{(x,z)} {=} i_{(x,z)} {+} i_{(\square,z)}, \;
j_{(z,x)} {=} i_{(z,x)} {+} i_{(z,\square)}, \;
j_{(u,v)} {=} i_{(u,v)} \mbox{ o.w.}
$$
Notations may seem heavy, but what they really express is that
replacing $\square$ by $x$ in $x^2\square^3z^2$ yields $x^5z^2$.

We now define an interpretation $P_.^y$ of indices containing
$\square$:
\begin{align*}
P^y_{\square z}(\state) &=
  \left| [\sigma(x){+}\sigma(y), \sigma(z)] \,\triangle\,
         [\sigma(x), \sigma(z)] \right| \\
P^y_{z \square}(\state) &=
  \left| [\sigma(z), \sigma(x){+}\sigma(y)] \,\triangle\,
         [\sigma(z), \sigma(x)] \right| \\
P^y_{\square \square}(\state) &= P_{x \square}(\state)
  = P_{\square x}(\state) = 0 \\
P^y_{uv}(\state) &= |[\sigma(u), \sigma(v)]| \mbox{ otherwise},
\end{align*}
we use $A \,\triangle\, B$ to denote the symmetric difference
of $A$ and $B$.
The quantity $P^y_{\square z}(\state)$ is the size change of
the interval $[x, z]$.  If $\state$ is the state before the
assignment and $\state'$ is the state after the assignment, we
have
$P_{xz}(\state) = P^y_{xz}(\state') + P^y_{\square z}(\state')$
and
$P_{zx}(\state) = P^y_{zx}(\state') - P^y_{z \square}(\state')$.
With these definitions, we have the following lemma.
\begin{lemma}
If $\state \rightarrow_C \state'$ and $\state(y)>0$ then
$P_Q(\state) = P^y_{\qsh^+_x Q}(\state')$.
\end{lemma}
\begin{proof}
\small
\begin{align*}
&P^y_{\qsh_x^+ Q}(\state')
= \sum_I q'_I \prod_{(u,v)} P_{uv}^{uv} \\
&= \sum_I q_{I[\square\gets x]} X_I \prod_z
  \binom{\square z {+} xz}{\square z}P_{\square z}^{\square z} P_{xz}^{xz}
  \binom{z\! \square\! {+} zx}{z \square}(-P_{z \square})^{z \square} P_{zx}^{zx} \\
&= \sum_J q_J X_J \sum_{I[\square\gets x] = J} \prod_z
  \binom{\square z {+} xz}{\square z}P_{\square z}^{\square z} P_{xz}^{xz}
  \binom{z\! \square\! {+} zx}{z \square}(-P_{z \square})^{z \square} P_{zx}^{zx} \\
&= \sum_J q_J X_J \prod_z
  \left( \sum_{\square z + xz = j_{xz}}
    \binom{\square z {+} xz}{\square z}P_{\square z}^{\square z} P_{xz}^{xz} \right)
  \left( \sum_{z\square {+} zx = j_{zx}} \dots \right) \\
&= \sum_J q_J X_J \prod_z
  (P_{xz} + P_{\square z})^{j_{xz}}
  (P_{zx} - P_{z \square})^{j_{zx}} \\
&= \sum_J q_J P_J(\state) = P_Q(\state)
\end{align*}
\end{proof}

@@

Given an index I from $\ind(V\cup \{ \square \})$, we define its projection
$J = I|_{0y} \in \ind(V)$ over the interval $[0,y]$ as
$$
j_{0y} = i_{0y} + \sum_{z \in V} i_{\square z} + i_{z \square}, \quad
j_{uv} = i_{uv} \mbox{ o.w.}
$$
Once again this is a degree preserving transformation.  We use it to
relate indices containing $P_{0y}^k$ and its possible \emph{developments}
in terms of $P_{\square z}$ and $P_{z \square}$.  A development gives a
possible rewrite of indices that contain the base polynomial $
|[0,y]|^k$ where k is non-zero.  Formally, it is a pair $\delta =
(\delta_0, \delta_m)$ of a natural number $\delta_0$ and a map $\delta_m$
from program variables to natural numbers such that $\delta_m(x) = 0$.
A development $\delta$ has degree $\delta_0 + \sum_z \delta_m(z)$.  We
will write $\Delta(J)$ for the set of all developments that have degree
$j_{0y}$.  We will also use the shorter functional notation $\delta(z)$
to denote $\delta_m(z)$.

To see the action of a development on an index, consider the
index $J = |[0,y]|^2$, the development $\delta = (2, \_ \mapsto 0)$
is in $\Delta(J)$ and transforms $J$ into itself.  The development
$\delta' = (1, \{ z \mapsto 1, \_ \mapsto 0\})$ is also into $\Delta(J)$ and
transforms $J$ into the resource polynomial
$|[0,y]|P_{\square z} + |[0,y]|P_{z \square}$.

If $Q$ is a resource polynomial on the index set $\ind(V\cup \{\square\})$ we
can now define the projection $Q' = Q|_{0y} \in \ind(V)$ as
$$
q'_J = \sum_{\delta \in \Delta(J)} q^\delta_J,
$$
where the numbers $q^\delta_J$ are constrained to satisfy the following
family of equations
$$
q_I = q^\delta_{I|_{0y}} \prod_z \binom{i_{\square z} + i_{z \square}}{i_{\square z}}.
$$

\begin{lemma}
If $Q \in \ind(V \cup \{\square\})$, then for all $\state$,
$P^y_Q(\state) = P_{Q|_{0y}}(\state)$.
\end{lemma}
\begin{proof}
\begin{align*}
&P_{Q|_{0y}}(\state)
  = \sum_J q_J P_J(\state)
  = \sum_J X_J \sum_{\delta \in \Delta(J)} q_J^\delta P_{0y}^{0y} \\
&= \sum_J X_J \sum_{\delta \in \Delta(J)}
    q_J^\delta P_{0y}^{\delta_0}
    \prod_z (P_{\square z} + P_{z \square})^{\delta(z)} \\
&= \sum_J X_J \sum_{\delta \in \Delta(J)}
    q_J^\delta P_{0y}^{\delta_0}
    \prod_z \sum_{i+j=\delta(z)} \binom{\delta(z)}{i}
      P_{\square z}^i P_{z \square}^j \\
&= \sum_J \sum_{I|_{0y} = J}
    q_J^{\delta_I} X_J P_{0y}^{i_{0y}}
    \prod_z \binom{i_{\square z} + i_{z \square}}{i_{\square z}}
      P_{\square z}^{i_{\square z}} P_{z \square}^{i_{z \square}} \\
&= \sum_I
    \left( q_{I|_{0y}}^{\delta_I}
    \prod_z \binom{i_{\square z} + i_{z \square}}{i_{\square z}}
    \right) P_I^y
  = \sum_I q_I P_I^y = P^y_Q(\state).
\end{align*}
\end{proof}
